---
output:
  word_document: default
  html_document: default
---
#Random Forest Assignment

```{r, include=FALSE}
library(tidyverse)
library(tidymodels)
library(caret)
library(gridExtra)
library(vip)
library(ranger)
library(naniar)
library(ggplot2)
```

```{r}
drug <- read_csv("drug_data-1.csv")
```

```{r}
names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity",
"Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive",
"SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis",
"Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh",
"LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")
str(drug)
```

```{r}
drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"
```

```{r}
drug_clean = drug %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44",
"45_54", "55_64", "65_"))) %>%
mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
mutate(Education = factor(Education, labels =
c("Under16", "At16", "At17", "At18", "SomeCollege",
"ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
mutate(Country = factor(Country,
labels = c("USA", "NewZealand", "Other", "Australia",
"Ireland","Canada","UK"))) %>%
mutate(Ethnicity = factor(Ethnicity,
labels = c("Black", "Asian", "White", "White/Black", "Other",
"White/Asian", "Black/Asian"))) %>% mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
select(-ID)
```
```{r, include=FALSE}
str(drug_clean)
```

```{r}
drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>% select(!(Semer:VSA))
names(drug_clean)
```

#Task 1

```{r}
gg_miss_var(drug_clean)
gg_miss_case(drug_clean)
vis_miss(drug_clean)
```

##Looks like no values are missing from the data set. 


#Task 2
```{r}
set.seed(1234)
drug_split = initial_split(drug_clean, prob = 0.70, strata = Nicotine)
train = training(drug_split)
test = testing(drug_split)
```

#Task 3
```{r}
p1 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill")
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill")
grid.arrange(p1,p2,p3,p4)


```

##Gender, Age, Education, and Country all seem to influence Nicotine. 

```{r}
p1 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Nicotine, y = Nscore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Escore)) + geom_boxplot()
p4 = ggplot(train, aes(x = Nicotine, y = Oscore)) + geom_boxplot()
  
grid.arrange(p1,p2,p3,p4)

```

##Ethnicity seems to have an effect. Nscore, Escore, and Oscore don't seem to change much. 

```{r}
p1 = ggplot(train, aes(x = Nicotine, y = Ascore)) + geom_boxplot()
p2 = ggplot(train, aes(x = Nicotine, y = Cscore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Impulsive)) + geom_boxplot()
p4 = ggplot(train, aes(x = Nicotine, y = SS)) + geom_boxplot()

grid.arrange(p1,p2,p3,p4)

```
##Cscore, Impulsive, and SS look good for correlation as well. 

#Task 4

```{r}
set.seed(123)
rf_folds = vfold_cv(train, v = 5)
```


```{r}
drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% 
  set_engine("ranger", importance = "permutation") %>% 
  set_mode("classification")

drug_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(drug_recipe)

set.seed(123)
rf_res = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = 10
)

```
```{r}
rf_grid = grid_regular(
  mtry(range = c(2, 8)), 
  min_n(range = c(5, 20)), 
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = rf_grid 
)
```

```{r}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  mutate(min_n = factor(min_n)) %>%
  ggplot(aes(mtry, mean, color = min_n)) +
  geom_line(alpha = 0.5, size = 1.5) +
  geom_point() +
  labs(y = "Accuracy")

```

```{r}
rf_grid = grid_regular(
  mtry(range = c(2, 8)), 
  min_n(range = c(5, 20)), 
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = rf_grid 
)
```



```{r}
best_rf = select_best(rf_res_tuned, "accuracy")

final_rf = finalize_workflow(
  drug_wflow,
  best_rf
)

final_rf
```


```{r}
final_rf_fit = fit(final_rf, train)

final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```

```{r}
final_rf_fit2 = fit(final_rf, test)

final_rf_fit2 %>% pull_workflow_fit() %>% vip(geom = "point")

```

##Importance across the variables shows that SS is the most influential variable among all others. 

##This model could be used in the real world! Sensation seeking associated with nicotine use could be used to predict how someone reacts with/to nicotine. An example of sensation seeking could be inability to focus without a cigarette/vape/dip or maybe used to predict what a nicotine user might use to replace the sensation of the nicotine device while trying to quit. 